---
title: "Little Loch Broom seabed map"
output: html_document
date: "2025-07-03"
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


rm(list = ls())

# Load required libraries
library(sf)
library(MESS)
library(RColorBrewer)
library(mgcv)
library(fields)
library(terra)
library(here)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(prettymapr)

op <- par(no.readonly = TRUE)
here::i_am("Sharing_code.Rmd")


######################################################### Load Bathymetry and Survey Data
# Load bathymetry raster and define heatmap palette for depth
depth <- rast(here("GIS_dat/WRoss_bathy.asc"))
plot(depth, main = "Bathymetry")

depth.heat <- colorRampPalette(rev(brewer.pal(9, 'Blues')[2:9]))

# Load survey data
dat22 <- read.csv(here("cam_data/Wester_Ross_2022data.csv"))
head(dat22)
str(dat22)
names(dat22)

######################################################### Subset and Recode Substratum Data

sub <- subset(dat22, Substratum1 != "")
sub <- droplevels(sub)

######################################################### Create Spatial Object for Camera Drops

# Convert substratum columns to factors and set factor levels consistently
sub$Substratum1 <- as.factor(sub$Substratum1)
sub$Substratum2 <- as.factor(sub$Substratum2)
sub$Substratum3 <- as.factor(sub$Substratum3)

sub$Substratum1 <- factor(sub$Substratum1, 
                          levels = c("mud", "muddy_sand", "sand", "gravel", "cob_boulder", "rock"))
sub$Substratum2 <- factor(sub$Substratum2, 
                          levels = c("mud", "muddy_sand", "sand", "gravel", "cob_boulder"))
sub$Substratum3 <- factor(sub$Substratum3, 
                          levels = c("muddy_sand", "sand", "gravel", "cob_boulder"))

# Convert data to sf object (WGS84)
sub.sf <- st_as_sf(sub, 
                   coords = c("Longitude", "Latitude"), remove = FALSE,
                   crs = "EPSG:4326")

# Reproject to British National Grid
sub.sf <- st_transform(sub.sf, crs = "EPSG:27700")

######################################################### Load Land and Shoreline Polygons

land <- st_read(here("GIS_dat/boundaryline/scotland_and_wales_region.shp"))
shore <- st_read(here("GIS_dat/boundaryline/high_water_polyline.shp"))

# Define bounding box for Little Loch Broom area
Wr.ext <- c(xmin = 186003.5, xmax = 209396.3, ymin = 886992.6, ymax = 899905.2)
Wr.bound <- c(xmin = 186010, xmax = 209396.3, ymin = 886992.6, ymax = 899893)

# Crop shapefiles to study area
shore.wr <- st_crop(shore, Wr.ext)
land.wr  <- st_crop(land, Wr.ext)

```


## Load 2023 Data
```{r load-2023-data}
# Import and prepare 2023 survey data
dat23 <- read.csv(here("cam_data/Wester_Ross_2023data.csv"))


######################################################### Subset and Recode Substratum Data

# Subset to records with Substratum1 classified
sub23 <- dat23 %>%
  filter(dat23$Substratum1 != "") %>%
  droplevels()

# Convert substratum to factors
sub23$Substratum1 <- factor(sub23$Substratum1,
                            levels = c("mud", "muddy_sand", "sand", "gravel", "cob_boulder", "rock"))
sub23$Substratum2 <- factor(sub23$Substratum2,
                            levels = c("mud", "muddy_sand", "sand", "gravel", "cob_boulder"))
sub23$Substratum3 <- factor(sub23$Substratum3,
                            levels = c("muddy_sand", "cob_boulder"))

######################################################### Convert to Spatial Object

# Convert to sf object (WGS84)
sub.sf23 <- st_as_sf(sub23, 
                     coords = c("Longitude", "Latitude"), remove = FALSE,
                     crs = "EPSG:4326")

# Reproject to British National Grid
sub.sf23 <- st_transform(sub.sf23, crs = "EPSG:27700")

######################################################### Add National Grid Coordinates Back to Dataframe

sub <- as.data.frame(sub.sf)
sub23 <- as.data.frame(sub.sf23)

######################################################### Define Colour Palettes for Prediction Maps


couls <- rev(brewer.pal(11, "Spectral"))
couls <- c("#FFFFFF", couls)
prob.heat <- colorRampPalette(couls)

```

## Extract depths and add to survey data
```{r extract-depths}
# Extract depth for 2022
depths <- extract(depth, sub.sf)

names(depths) <- c("ID", "Depth")

# Add depth values to spatial and data frame objects
sub.sf <- cbind(sub.sf, Depth = depths$Depth)
sub    <- cbind(sub,    Depth = depths$Depth)

```


# functions
```{r functions}
# Function to plot base map with depth, land and rectangle
plot_base_map <- function(depth, land, Wr.bound, depth_palette) {
  par(mar = c(0, 0, 0, 0) + 0.1)
  plot(st_geometry(land), col = NA, border = NA)
  rect(Wr.bound["xmin"], Wr.bound["ymin"], Wr.bound["xmax"], Wr.bound["ymax"], 
       col = "lightsteelblue4", border = NA)
  plot(st_geometry(land), add = TRUE, col = 'grey40', border = 'grey60')
  plot(depth, add = TRUE, col = depth_palette(1000), legend = FALSE, axes = FALSE)
}


# Function to plot base map with no depth
plot_land <- function( land, Wr.bound) {
  par(mar = c(0, 0, 0, 0) + 0.1)
  plot(st_geometry(land), col = NA, border = NA)
  rect(Wr.bound["xmin"], Wr.bound["ymin"], Wr.bound["xmax"], Wr.bound["ymax"], 
       col = "lightsteelblue4", border = NA)
  plot(st_geometry(land), add = TRUE, col = 'grey40', border = 'grey60')
  
}


# Function to plot substratum points for a given sf object
plot_sub_points <- function(sub_sf, substratum_cols) {
  plot(st_geometry(sub_sf), pch = 19, cex = 1.5, add = TRUE,
       col = substratum_cols[as.character(sub_sf$Substratum1)])
  plot(st_geometry(sub_sf), pch = 19, cex = 1, add = TRUE,
       col = substratum_cols[as.character(sub_sf$Substratum2)])
  plot(st_geometry(sub_sf), pch = 19, cex = 0.5, add = TRUE,
       col = substratum_cols[as.character(sub_sf$Substratum3)])
}

# Function to plot substratum legend
plot_sub_legend <- function(substratum_cols) {
  par(mar = c(0, 0, 0, 0) + 0.1)
  plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
  legend('center',
         legend = names(substratum_cols),
         fill = substratum_cols,
         bg = 'white')
}

# Function to plot depth colour scale legend
plot_depth_legend <- function(depth, depth_palette) {
  par(mar = c(0, 0, 0, 0) + 0.1)
  image.plot((depth * -1),
             horizontal = FALSE,
             smallplot = c(0.6, 0.63, 0.1, 0.9),
             col = depth_palette(1000),
             legend.only = TRUE,
             axis.args = list(
               cex.axis = 1,
               at = c(0, 20, 40, 60, 80, 100, 120, 140, 160, 180),
               labels = rev(c("0", "20", "40", "60", "80", "100", "120", "140", "160", "180")),
               tcl = 0.3, lwd = 0.5, hadj = 0.5),
             zlim = c(0, 181),
             axes = FALSE)
}

# addalpha()
addalpha <- function(colors, alpha=1.0) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

# Define substratum colours
substratum_cols <- c(
  "mud" = rgb(255,127,0, maxColorValue = 255),
  "muddy_sand" = rgb(77,175,74, maxColorValue = 255),
  "sand" = rgb(255,255,51, maxColorValue = 255),
  "gravel" = rgb(228,26,28, maxColorValue = 255),
  "cob_boulder" = rgb(152,78,163, maxColorValue = 255),
  "rock" = rgb(55,126,184, maxColorValue = 255)
)

```

# Sample location maps
```{r sample-location-map}


# Get UK map from Natural Earth
uk <- ne_countries(scale = 10, country = "United Kingdom", returnclass = "sf")
eir<- ne_countries(scale = 10, country = "Ireland", returnclass = "sf")
# use one drop location for site location
site <- st_as_sf(sub23[1,], 
                     coords = c("Longitude", "Latitude"), remove = FALSE,
                     crs = "EPSG:4326")


point_col = c('darkblue')

# Set up plot window
par(mar = c(1, 1, 1, 1))  # margins: bottom, left, top, right
# Optional: save as PDF
#pdf(here("figures/1_site_locations.pdf"), width = 5, height = 5)

plot(st_geometry(uk), col = "grey80", border = "black")
plot(st_geometry(eir), col = "grey80", border = "black", add=T)


plot(st_geometry(site), col = point_col, pch = 18, cex = 2, add=TRUE)

addscalebar(plotepsg = 4326,  pos = "bottomright", widthhint =0.15)
addnortharrow(pos = "topright", scale = 0.8)

#dev.off()



# Optional: save as PDF
# pdf(here("figures/1_sample_locations22.pdf"), width = 5, height = 5)

# Set up plot margins
par(mar = c(0, 0, 0, 0) + 0.1)


# 
# plot(st_geometry(sub.sf), add = TRUE, pch = 19, col = "deeppink4", cex = 1)
# 
#
# plot(st_geometry(sub.sf23), add = TRUE, pch = 19, col = "blue4", cex = 1)

# Plot for 2022 data
plot_base_map(depth, land.wr, Wr.bound, depth.heat)
plot_sub_points(sub.sf, substratum_cols)

#dev.off()

# Optional: save as PDF
#pdf(here("figures/1_sample_locations23.pdf"), width = 5, height = 5)

# Plot for 2023 data
plot_base_map(depth, land.wr, Wr.bound, depth.heat)
plot_sub_points(sub.sf23, substratum_cols)


#dev.off()

plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
# Optional: save as PDF
#pdf(here("figures/1_substrata_leg.pdf"), width = 5, height = 5)
# Substratum legend

plot_sub_legend(substratum_cols)

# dev.off()
 
 
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
# Optional: save as PDF
#pdf(here("figures/1_depth_leg.pdf"), width = 5, height = 5)

# Depth colour scale legend

plot_depth_legend(depth, depth.heat)


# dev.off()

par(op)

```

# Models

## Mud
```{r}

#For this we will not use any test/train partition but the entire dataset:

# refit model to all
head(sub)

GAM1 <- gam(formula = mud ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)
summary(GAM1)

## make data frame for predictions


depth_wgs<- project(depth, 'EPSG:4326')
depth.val<- values(depth_wgs, dataframe=T)
depth.val<- as.data.frame(depth_wgs, cell=TRUE, xy=TRUE)
head(depth.val)
summary(depth.val)
names(depth.val)[c(2,3,4)]<- c('Longitude', 'Latitude', 'Depth')


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)
head(PredGAM)
summary(PredGAM)

mud_wgs <- rast(depth_wgs)
mud_wgs[PredGAM$cell] <- PredGAM[,5]


mud.r<- project(mud_wgs, 'EPSG:27700')


# Optional: save as PDF
#pdf(here("figures/2_mud_prediction.pdf"), width = 5, height = 5)
par(mar = c(0, 0, 0, 0) + 0.1)

plot_land(land.wr, Wr.bound, depth.heat)



plot(mud.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()


```

## model plots legend
```{r}
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
# Optional: save as PDF
#pdf(here("figures/2_prediction_leg.pdf"), width = 5, height = 5)

image.plot((mud.r), horizontal=T, smallplot= c(0.1, 0.9,0.6,0.63), col = prob.heat(1000), legend.only=TRUE,  axis.args = list(cex.axis = 1,
at = c(0, 50, 100), labels =c("0","0.5","1.0"), 
tcl=0.3, lwd=0.5, hadj=0.5), axes=F, zlim=c(0, 100))

# dev.off()

```

## Muddy_sand
```{r}

### GAM
#For this we will not use any test/train partition but the entire dataset:

# refit model to all

GAM1 <- gam(formula = muddy_sand ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)
head(PredGAM)
summary(PredGAM)

muddyS_wgs <- rast(depth_wgs)
muddyS_wgs[PredGAM$cell] <- PredGAM[,5]

muddyS.r<- project(muddyS_wgs, 'EPSG:27700')


# Optional: save as PDF
#pdf(here("figures/2_muddyS_prediction.pdf"), width = 5, height = 5)

plot_land(land.wr, Wr.bound, depth.heat)

plot(muddyS.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()


```

## sand
```{r}

### GAM
#For this we will not use any test/train partition but the entire dataset:

# refit model to all

GAM1 <- gam(formula = sand ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)


sand_wgs <- rast(depth_wgs)
sand_wgs[PredGAM$cell] <- PredGAM[,5]

sand.r<- project(sand_wgs, 'EPSG:27700')
# Optional: save as PDF
#pdf(here("figures/2_sand_prediction.pdf"), width = 5, height = 5)

plot_land(land.wr, Wr.bound, depth.heat)


plot(sand.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()

```


## gravel 
```{r}

### GAM
#For this we will not use any test/train partition but the entire dataset:

# refit model to all

GAM1 <- gam(formula = gravel ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)
head(PredGAM)
summary(PredGAM)

gravel_wgs <- rast(depth_wgs)
gravel_wgs[PredGAM$cell] <- PredGAM[,5]

gravel.r<- project(gravel_wgs, 'EPSG:27700')
# Optional: save as PDF
#pdf(here("figures/2_gravel_prediction.pdf"), width = 5, height = 5)

plot_land(land.wr, Wr.bound, depth.heat)



plot(gravel.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()

```




## Cob_boulder 
```{r}

### GAM
#For this we will not use any test/train partition but the entire dataset:

# refit model to all

GAM1 <- gam(formula = cob_boulder ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)
head(PredGAM)
summary(PredGAM)

cob_wgs <- rast(depth_wgs)
cob_wgs[PredGAM$cell] <- PredGAM[,5]

cob.r<- project(cob_wgs, 'EPSG:27700')

# Optional: save as PDF
#pdf(here("figures/2_cob_prediction.pdf"), width = 5, height = 5)

plot_land(land.wr, Wr.bound, depth.heat)


plot(cob.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()


```


## rock 
```{r}

### GAM
#For this we will not use any test/train partition but the entire dataset:

# refit model to all

GAM1 <- gam(formula = rock ~  s(Longitude, Latitude)+ s(Depth), data = sub, family = binomial)


PredGAM <- as.data.frame(predict(GAM1, newdata=depth.val, type="response", se.fit=T))
PredGAM<- cbind(depth.val, PredGAM)
head(PredGAM)
summary(PredGAM)

rock_wgs <- rast(depth_wgs)
rock_wgs[PredGAM$cell] <- PredGAM[,5]

rock.r<- project(rock_wgs, 'EPSG:27700')

# Optional: save as PDF
#pdf(here("figures/2_rock_prediction.pdf"), width = 5, height = 5)

plot_land(land.wr, Wr.bound, depth.heat)


plot(rock.r, add=T, legend=FALSE, axes=FALSE, col=prob.heat(1000) )

# dev.off()



```

# K-fold cross validation
## Code provided to split data into training and test sets for five folds across 100 CV iterations. Used to test the model "in-bag" prediction accuracy.
```{r}



K <- 5
N <- 100
first.run <- TRUE

# k-fold splitting function
kfoldSplit <- function(x, k=K, train=TRUE){
  x <- sample(x, size = length(x), replace = FALSE)
  out <- suppressWarnings(split(x, factor(1:k)))
  if(train) out <- lapply(out, function(x, len) (1:len)[-x], len=length(unlist(out)))
  return(out)
}

kfolds <- kfoldSplit(1:nrow(sub), k = K, train = TRUE)

# Sediment types
sediments <- c("mud", "muddy_sand", "sand", "gravel", "cob_boulder", "rock")

# Initialise list to store RMSE matrices per seabed
evalList <- lapply(sediments, function(x) matrix(NA, nrow = K, ncol = 1, dimnames = list(1:K, "GAM")))
names(evalList) <- sediments

for (run in 1:N) {
  
  for (i in 1:K) {
    
    cat("Run", run, "- K-fold", i, "of", K, "...\n")
    
    idx <- kfolds[[i]]
    idxBool <- (1:nrow(sub)) %in% idx
    
    for (sed in sediments) {
      # Observed test values
      obs.test <- sub[!idxBool, sed]
      
      # Fit GAM on training data
      GAM <- gam(
        formula = as.formula(paste(sed, "~ s(Longitude, Latitude) + s(Depth)")),
        data = sub.sf[idx, ],
        family = binomial
      )
      
      # Predict on test data
      pred.test <- predict(GAM, newdata = sub.sf[!idxBool, ], type = "response")
      
      # Calculate RMSE and store
      rmse <- sqrt(mean((pred.test - obs.test)^2))
      evalList[[sed]][i, "GAM"] <- rmse
    }
    
  }
  
  # Combine evaluation results for all sediments
  evals.tmp <- do.call(rbind, evalList)
  
  if (first.run) {
    evals <- evals.tmp
    first.run <- FALSE
  } else {
    evals <- rbind(evals, evals.tmp)
  }
  
  print(evals)
}


round(apply(evals,2,FUN = function(x, na.rm=TRUE) c(mean(x,na.rm=TRUE), sd(x, na.rm=TRUE))),3)


```


# accuracy curve across probabilities
```{r}
par(op)

# Clean Substratum columns
True.dat <- sub.sf23
substratum.cols <- c("Substratum1", "Substratum2", "Substratum3")


# convert NAs to "No" in Substratum columns
for (col in substratum.cols) {
  True.dat[[col]] <- as.character(True.dat[[col]])
  True.dat[[col]][is.na(True.dat[[col]])] <- "No"
}



# Substratum labels and raster stack
substrata <- c("mud", "muddy_sand", "sand", "gravel", "cob_boulder", "rock")
sub.stack <- rast(list(mud.r, muddyS.r, sand.r, gravel.r, cob.r, rock.r))
sub.trim <- classify(sub.stack, cbind(0, NA))  # set 0 to NA

# Build truth dataframe: Yes/No for each substratum type
truth <- sapply(substrata, function(sed) {
  apply(True.dat[, substratum.cols], 1, function(row) {
    if (any(trimws(row) == sed)) "Yes" else "No"
  })
})
colnames(truth) <- substrata
truth.sp <- sub.sf23

# Probability thresholds
p.vals <- seq(0, 1, by = 0.01)

# Initialise list to hold accuracy values
overall.acc <- matrix(NA, nrow = length(p.vals), ncol = length(substrata))
colnames(overall.acc) <- substrata

# Loop through substratum layers
for (k in seq_along(substrata)) {
  cat("Processing:", substrata[k], "\n")
  
  layer.trim <- sub.trim[[k]]
  
  acc.vals <- numeric(length(p.vals))
  
  for (i in seq_along(p.vals)) {
    # Apply threshold
    pred.layer <- classify(layer.trim, cbind(-Inf, p.vals[i], NA))
    
    # Extract prediction values at sample points
    pred.prob <- extract(pred.layer, truth.sp)
    
    # Convert to binary: 1 if non-NA, 0 if NA
    pred.bin <- ifelse(!is.na(pred.prob[,2]), "Yes", "No")
    
    # Accuracy: proportion correct
    acc <- mean(pred.bin == truth[,k], na.rm = TRUE)
    acc.vals[i] <- acc
  }
  
  overall.acc[,k] <- acc.vals
}

# Mean and CI across substrata (per threshold)
acc.comb <- rowMeans(overall.acc)
acc.sd <- apply(overall.acc, 1, sd)
acc.up <- acc.comb + (acc.sd / sqrt(length(substrata)) * 1.96)
acc.lo <- acc.comb - (acc.sd / sqrt(length(substrata)) * 1.96)

# Plot results
# Optional: save as PDF
#pdf(here("figures/4_Precision_vs_Accuracy.pdf"), width = 5, height = 5)

prec <- rev(p.vals)

plot(prec, acc.comb,
     type='l', ylim=c(0,1), lwd=3, xlim=c(0,1.1),
     xlab='Precision', ylab='Accuracy')

# Find optimum point on mean line for all seabed types
dist <- sqrt((1 - prec)^2 + (1 - acc.comb)^2)
idx <- which.min(dist)

# Add point
points(prec[idx],
       acc.comb[idx],
       pch=18,
       cex=2,
       col="black")

for (j in 1:ncol(overall.acc)) {

    lines(prec,
          overall.acc[,j],
          col=adjustcolor(test.cols[j], alpha.f=0.5),
          lwd=2)

    # Find optimum point for these lines
    dist <- sqrt((1 - prec)^2 +
                 (1 - overall.acc[,j])^2)
    idx <- which.min(dist)

    points(prec[idx],
           overall.acc[idx,j],
           pch=18,
           cex=2,
           col=adjustcolor(test.cols[j], alpha.f=0.5))
}

lines(prec, acc.comb,
     lwd=3 )
lines(prec, acc.up, lty=2, lwd=2, col='grey40')
lines(prec, acc.lo, lty=2, lwd=2, col='grey40')

#dev.off()


# Add overall max point
best.overall <- which.max(acc.comb)
points(p.vals[best.overall], acc.comb[best.overall], pch=18, cex=3)
text(p.vals[best.overall]+0.05, acc.comb[best.overall]+0.05, labels=round(acc.comb[best.overall], 2))

legend('bottomright', legend=substrata, fill=substratum_cols, bg='white')

######################################################### Table for working out the optimal values
line.names <- c("Combined",
                "mud",
                "muddy-sand",
                "sand",
                "gravel",
                "cobble-boulder",
                "rock")

# Main (black) line
dist <- sqrt((1 - prec)^2 + (1 - acc.comb)^2)
idx <- which.min(dist)

opt.points <- data.frame(
  Line = line.names[1],
  Precision = prec[idx],
  Accuracy = acc.comb[idx],
  Distance = dist[idx]
)

# Individual habitat lines
for (j in 1:ncol(overall.acc)) {

  dist <- sqrt((1 - prec)^2 + (1 - overall.acc[,j])^2)
  idx <- which.min(dist)

  opt.points <- rbind(
    opt.points,
    data.frame(
      Line = line.names[j + 1],
      Precision = prec[idx],
      Accuracy = overall.acc[idx, j],
      Distance = dist[idx]
    )
  )
}

# Round values for display
opt.points$Precision <- round(opt.points$Precision, 3)
opt.points$Accuracy  <- round(opt.points$Accuracy, 3)
opt.points$Distance  <- round(opt.points$Distance, 3)

print(opt.points)





# AUC calculations
MESS::auc(p.vals, acc.comb, type = "linear")

```




# combined map with informed thresholds
```{r combined map}

slope1.thresholds <- c(1-opt.points$Precision[-1])  # example values from above table
# or you could use another informed, sensible value. For example the global average from the black line in the plot above

# Name rasters and matching thresholds in a named list
substratum_rasters <- list(
  mud = mud.r,
  muddy_sand = muddyS.r,
  sand = sand.r,
  gravel = gravel.r,
  cob_boulder = cob.r,
  rock = rock.r
)

# Apply optimal threshold trimming
substratum_trimmed <- mapply(function(raster, thresh) {
  raster[raster < thresh] <- NA
  return(raster)
}, substratum_rasters, slope1.thresholds, SIMPLIFY = FALSE)

# Now assign each back to a clean named object
mud.trim         <- substratum_trimmed$mud
muddy_sand.trim  <- substratum_trimmed$muddy_sand
sand.trim        <- substratum_trimmed$sand
gravel.trim      <- substratum_trimmed$gravel
cob_boulder.trim <- substratum_trimmed$cob_boulder
rock.trim        <- substratum_trimmed$rock



test.cols <- c(
     rgb(253,192,134, maxColorValue = 255),
     rgb(127,201,127, maxColorValue = 255),
     rgb(255,255,153, maxColorValue = 255),
     rgb(240,2,127, maxColorValue = 255),
     rgb(190,174,212, maxColorValue = 255),
     rgb(56,108,176, maxColorValue = 255)
     )

mud.alt <- colorRampPalette(c('wheat', test.cols[1], substratum_cols[1]), alpha=TRUE, bias=1)

mud.s.alt <- colorRampPalette(c('darkseagreen1', test.cols[2], substratum_cols[2]), alpha=TRUE, bias=1)
sand.alt <- colorRampPalette(c('lightgoldenrodyellow', test.cols[3], substratum_cols[3]), alpha=TRUE, bias=1.5)
gravel.alt <- colorRampPalette(c('lightpink1', test.cols[4], substratum_cols[4]), alpha=TRUE, bias=0.3)
cob_bol.alt<- colorRampPalette(c('thistle1', test.cols[5], substratum_cols[5]), alpha=TRUE, bias=1.5)
rock.alt<- colorRampPalette(c('lightsteelblue2', test.cols[6], substratum_cols[6]), alpha=TRUE, bias=1)



 


# Optional: save as PDF
#pdf(here("figures/5_combined_prediction.pdf"), width = 5, height = 5)
 
 plot_land(land.wr, Wr.bound, depth.heat)

 
 
plot(cob_boulder.trim, add=T, legend=FALSE, axes=FALSE, col=cob_bol.alt(1000), alpha=0.7)

plot(gravel.trim, add=T, legend=FALSE, axes=FALSE, col=gravel.alt(1000), alpha=0.4)

plot(sand.trim, add=T, legend=FALSE, axes=FALSE, col=sand.alt(1000), alpha=0.6)


plot(muddy_sand.trim, add=T, legend=FALSE, axes=FALSE, col=mud.s.alt(1000), alpha=0.5)


plot(rock.trim, add=T, legend=FALSE, axes=FALSE, col=rock.alt(1000), alpha=0.8)


plot(mud.trim, add=T, legend=FALSE, axes=FALSE, col=mud.alt(1000), alpha=0.6)


plot(st_geometry(sub.sf23), pch=19, add=T, col=
       substratum_cols[sub.sf23$Substratum1], cex=1.3)

plot(st_geometry(sub.sf23), pch=19, add=T, col=
       substratum_cols[sub.sf23$Substratum2], cex=0.8)

plot(st_geometry(sub.sf23), pch=19, add=T, col=
       substratum_cols[c(2,3,4,5)][sub.sf23$Substratum3], cex=0.5)

# dev.off()


```

# combined plot legend
```{r}
mud.alt.single <- addalpha(substratum_cols[1], alpha=0.7)
mud.s.alt.single <- addalpha(substratum_cols[2], alpha=0.7)
sand.alt.single <- addalpha(substratum_cols[3], alpha=0.7)
gravel.alt.single <-addalpha(substratum_cols[4], alpha=0.7)
cob_bol.alt.single<- addalpha(substratum_cols[5], alpha=0.7)
rock.alt.single<- addalpha(substratum_cols[6], alpha=0.7)

leg<- c("mud", "muddy sand", "sand", "gravel", "cobble boulder", "rock" )

plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)


 leg.cols<- c(mud.alt.single, mud.s.alt.single, sand.alt.single,
gravel.alt.single, cob_bol.alt.single, rock.alt.single)


# Optional: save as PDF
#pdf(here("figures/5_legend.pdf"), width = 5, height = 5)

par(mar=c(0,0,0,0)+0.1)

legend('center', bg='white', legend=leg, fill= leg.cols)

#dev.off()


```

